---
title: "Zebrafish"
author: "Felix Haidle"
output:
  BiocStyle::html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The Zebrafish data set is part of the research on the effect of USP39 knockouts on splicing expression.


# Raw Data Procession

Raw data was deposited in the GEO database with the accession number:


## Trimming of raw data
```{bash,eval=FALSE}
java -jar \
        /share/project/zarnack/felix/tools/Trimmomatic/Trimmomatic/dist/jar/trimmomatic-0.40-rc1.jar \
        SE \
        $file \
        ./$outfile \
        ILLUMINACLIP:/share/project/zarnack/felix/tools/Trimmomatic/Trimmomatic/adapters/TruSeq3-SE.fa:2:30:10
```

## Read Mapping

RNAStar has been used to align the reads to the reference genome
the reference genome used was the GRCz11 version 110

command structure for this dataset:
```{bash,eval=FALSE}
$path/tools/STAR/bin/Linux_x86_64_static/STAR \ 
--runThreadN 8 \
--genomeDir $path/genomes/GRCz11_rl112 \
--readFilesIn $file \
--readFilesCommand zcat \
--outSAMtype BAM SortedByCoordinate \
--outFilterMultimapNmax 1 \
--outFilterMismatchNmax 999 \
--outFilterMismatchNoverLmax 0.04  \
--outFileNamePrefix starout/"${substring}" 

```



## Splicing analysis

"MAJIQ" was used to identify splicing events. Note that we removed one of the samples following quality control.

Build command:
```{bash,eval=FALSE}
GFF3="/share/project/zarnack/felix/USP39/zebrafish/HB_folder/genome/Danio_rerio.GRCz11.110.gff3"
CONFIGFILE="./majiq/majiq.config"
BUILDDIR="./majiq/build"

majiq deltapsi \
-grp1 \
"${BUILDDIR}/trimmed_C_repA_S1_L001_R1_001Aligned.sortedByCoord.out.majiq" \
"${BUILDDIR}/trimmed_C_repB_S2_L001_R1_001Aligned.sortedByCoord.out.majiq" \
"${BUILDDIR}/trimmed_C_repC_S3_L001_R1_001Aligned.sortedByCoord.out.majiq" \
"${BUILDDIR}/trimmed_C_repD_S4_L001_R1_001Aligned.sortedByCoord.out.majiq" \
-grp2 \
"${BUILDDIR}/trimmed_M_repA_S5_L001_R1_001Aligned.sortedByCoord.out.majiq" \
"${BUILDDIR}/trimmed_M_repB_S6_L001_R1_001Aligned.sortedByCoord.out.majiq" \
"${BUILDDIR}/trimmed_M_repC_S7_L001_R1_001Aligned.sortedByCoord.out.majiq" \
-j $SLURM_CPUS_PER_TASK \
--output-type voila \
-o $DELTAPSIDIR \
-n C M   

```
the config file was: 

```{bash,eval=FALSE}
[info]
bamdirs=/share/project/zarnack/felix/USP39/zebrafish_from_fastq/HB_data/starout
genome=GRCz11
strandedness=forward
[experiments]
C=trimmed_C_repA_S1_L001_R1_001Aligned.sortedByCoord.out,trimmed_C_repB_S2_L001_R1_001Aligned.sortedByCoord.out,trimmed_C_repC_S3_L001_R1_001Aligned.sortedByCoord.out,trimmed_C_repD_S4_L001_R1_001Aligned.sortedByCoord.out
M=trimmed_M_repA_S5_L001_R1_001Aligned.sortedByCoord.out,trimmed_M_repB_S6_L001_R1_001Aligned.sortedByCoord.out,trimmed_M_repC_S7_L001_R1_001Aligned.sortedByCoord.out

```



to evaluate changings between groups "majiq deltapsi" was used

```{bash,eval=FALSE}
majiq deltapsi \
-grp1 \
"${BUILDDIR}/trimmed_C_repA_S1_L001_R1_001Aligned.sortedByCoord.out.majiq" \
"${BUILDDIR}/trimmed_C_repB_S2_L001_R1_001Aligned.sortedByCoord.out.majiq" \
"${BUILDDIR}/trimmed_C_repC_S3_L001_R1_001Aligned.sortedByCoord.out.majiq" \
"${BUILDDIR}/trimmed_C_repD_S4_L001_R1_001Aligned.sortedByCoord.out.majiq" \
-grp2 \
"${BUILDDIR}/trimmed_M_repA_S5_L001_R1_001Aligned.sortedByCoord.out.majiq" \
"${BUILDDIR}/trimmed_M_repB_S6_L001_R1_001Aligned.sortedByCoord.out.majiq" \
"${BUILDDIR}/trimmed_M_repC_S7_L001_R1_001Aligned.sortedByCoord.out.majiq" \
-j $SLURM_CPUS_PER_TASK \
--output-type voila \
-o $DELTAPSIDIR \
-n C M   
```




to identify specific event types modulize was used
```{bash,eval=FALSE}
MODULIZEDIR="./majiq/modulize"
SPLICEGRAPH="./majiq/build/splicegraph.sql"
VOILAFILES="./majiq/deltapsi/C-M.deltapsi.voila"

voila modulize \
--changing-between-group-dpsi 0.05 \
--non-changing-between-group-dpsi 0.02 \
--changing-between-group-dpsi-secondary 0.025 \
--probability-changing-threshold 0.9 \
--probability-non-changing-threshold 0.5 \
--show-all \
-j $SLURM_CPUS_PER_TASK \
-d $MODULIZEDIR \
$SPLICEGRAPH $VOILAFILES
```



# DESeq2 Analysis


## featureCounts
featureCounts was used to determine the genexpression of each sample

```{bash,eval=FALSE}
featureCounts \
-t exon \
-g gene_id \
-s 1 \
-a /share/project/zarnack/felix/USP39/zebrafish/HB_folder/genome/Danio_rerio.GRCz11.110.gtf \
-o counts.txt \
/share/project/zarnack/felix/USP39/zebrafish_from_fastq/HB_data/starout/*.bam
```


## Libraries
```{r libraries, warning=FALSE,message=FALSE,echo=TRUE}
library(DESeq2)
library(data.table)
library(dplyr)
library(openxlsx)
```


## DESeq calculations
```{r deseq2,warning=FALSE,message=FALSE}

rds_path_dds <- ("./data/rds_objects/Zebrafish_dds.rds")
if (!file.exists(rds_path_dds)) {
  # read in the featureCounts output
  feature_counts <- fread("./data/feature_counts/Zebrafish_counts.txt", header = TRUE, skip = 1)
  
  # extract the count columns
  sample_columns <- grepl(".bam",colnames(feature_counts))
  counts <- feature_counts[, ..sample_columns]
  
  counts <- as.data.frame(counts)
  
  # renaming columns to remove the path and file ending
  column_names <- tools::file_path_sans_ext(basename(colnames(counts)))
  # removing the Aligned.sortedByCoord.out file ending to improve readability
  new_column_names <- gsub("Aligned.sortedByCoord.out", "", column_names)
  colnames(counts) <- new_column_names
  
  
  # renaming rows to be the gene id
  rownames(counts) <- as.character(feature_counts$Geneid)
  
  
  # creating vector for sample numbering
  replicate_numbers <- rep(1:(ncol(counts)/2), length.out =ncol(counts))
  
  # creating condition vector 
  # !!!warning!!! needs to be adjustet for different datasets, depending on how the samples are arranged in the counts.txt
  condition_vector<- rep(c("wt", "ko"), each = ncol(counts)/2)
  
  
  
  matrix_counts <- as.matrix(counts)
  
  
  
  
  
  
  # sample info dataframe
  sample_info <- data.frame(sampleName = c(colnames(matrix_counts)),
                            condition = condition_vector,
                            replicate = replicate_numbers
                            )
  
  
  # creating dds object
  dds <- DESeqDataSetFromMatrix(countData = matrix_counts, colData = sample_info, design = ~condition, tidy = FALSE)
  
  # adding gene names to the dds rows
  rownames(dds) <- feature_counts$Geneid
  
  # releveling dds object
  dds$condition = relevel(dds$condition,ref = "wt")
  
  
  
  # DESeq analysis
  dds <- DESeq(dds)
  
  
  saveRDS(dds,rds_path_dds)
} else {
  # Load the dataframe from the RDS object
  dds <- readRDS(file = rds_path_dds)
}

sizeFactors(dds)


# setting an alpha value
alpha_value <- 0.05


rld <- rlog(dds, blind = FALSE)

resultsDds <- results(dds, contrast = c("condition","ko","wt"), alpha = alpha_value)


# adding shrunken log2FoldChange values
# calculating shrunken values
results_shrink <- lfcShrink(dds=dds, res= resultsDds, coef=2, type = "ashr")
resultsDds$log2FoldChange_shrink <- results_shrink$log2FoldChange


# turn resultsobject into a dataframe for export
results_dataframe <- as.data.frame(resultsDds)



# results datframe + counts counts(dds, normalised=T) + assay(rld)



# renaming the extra dataframe columns
# for the normalized counts
counts_dds_normalized <- counts(dds, normalized=TRUE) 
current_names <- colnames(counts_dds_normalized)
new_names <- paste0(current_names, "_normalized_counts")
colnames(counts_dds_normalized) <- new_names

# for the log transformed counts
counts_dds_log_transformed <- assay(rld) 
current_names <- colnames(counts_dds_log_transformed)
new_names <- paste0(current_names, "_log_transformed_counts")
colnames(counts_dds_log_transformed) <- new_names



# combining the different counts in one dataframe
big_results_table <- cbind(results_dataframe,
                           counts_dds_log_transformed,
                           counts_dds_normalized)


# output of the data to an excel sheet

# name of the output file
excel_file <- "data/DESeq2_results/Zebrafish_deseq2.xlsx"

# Write the dataframe to an Excel file
write.xlsx(big_results_table, excel_file, sheetName = "Sheet1", rowNames = TRUE)
```


# Session Info

```{r}
sessionInfo()
```